---
title: "Data checks"
author: "Lukáš Hejtmánek"
format: 
  html:
    self-contained: true
---

## Data checks

```{r}
#| echo: false
#| warning: false 
library(tidyverse)
library(here)
library(papaja)
library(correlation)
library(gt)
library(report)
here::i_am("reports/data-checks.qmd")
source(here("reports/elements/loading-preprocessing.R"))
source(here("reports/elements/analytical-help.R"))
load(here::here("temp/df_forest.RData"))
load(here::here("temp/df_health.RData"))
df_forest <- remove_unnecessary_columns(df_forest)
```

```{r processing }
#| echo: false
bad_participants <- get_participants_without_both_sessions(df_forest)
df_forest <- filter(df_forest, !ID %in% bad_participants)

df_measures_original <- format_important_measures(df_forest)
df_measures <- scale_measures(df_measures_original) %>%
  filter_measures()
```

```{r colnames}
colnames(df_forest)
```

## Overall report

```{r report}
report::report_table(df_forest) %>%
  select(-c(n_Obs, percentage_Missing, Skewness, Kurtosis, n_Entries)) %>%
  filter(!is.na(Mean)) %>%
  gt() %>%
    fmt_number(columns = where(~ is.numeric(.x)), decimals = 3)
```


### Checking session number

```{r effect of first_session}
df_forest <- add_first_condition(df_forest)

df_measures %>%
  left_join(select(df_forest, ID, condition, first_condition), by = c("ID", "condition")) %>%
  filter(condition != first_condition) %>%
  group_by(what, when) %>%
  group_modify(~report_table(t.test(.x$value ~ .x$first_condition)), .keep = FALSE) %>%
  select(-c(Parameter, Group, Method, Alternative, CI), -contains("CI")) %>%
  mutate(p = apa_p(p)) %>%
  gt() %>%
    tab_header(title = "Effect of first session") %>%
    fmt_number(columns = where(~ is.numeric(.x)), decimals = 3)
```


```{r session number}
df_measures %>%
  left_join(select(df_forest, ID, condition, session), by = c("ID", "condition")) %>%
  filter(when == "post") %>%
  group_by(what, condition) %>%
  group_modify(~report_table(t.test(.x$value ~ .x$session)), .keep = FALSE) %>%
  select(-c(Parameter, Group, Method, Alternative, CI), -contains("CI")) %>%
  mutate(p = apa_p(p)) %>%
  gt() %>%
    tab_header(title = "Effect of session number") %>%
    fmt_number(columns = where(~ is.numeric(.x)), decimals = 3)
```

### Checking administrator

```{r administrator effect descriptives}
df_forest %>%
  select(administrator, starts_with("relax"), starts_with("POMS"), starts_with("blood")) %>%
  group_by(administrator) %>%
  summarise_all(mean, na.rm = TRUE) %>%
  pivot_longer(-administrator, names_to = "variable", values_to = "value") %>%
  pivot_wider(names_from = "administrator", values_from = "value") %>%
  gt() %>%
    tab_header(title = "Effect of administrator") %>%
    fmt_number(columns = where(~ is.numeric(.x)), decimals = 3)
```

```{r administrator effect statistics}
df_measures %>%
  select(-z_value) %>%
  pivot_wider(names_from = "when", values_from = "value", id_cols = c(ID, condition, what, category)) %>%
  mutate(diff = post - pre) %>%
  pivot_longer(cols = c(pre, post, diff), names_to = "when", values_to = "value") %>%
  left_join(select(df_forest, ID, condition, administrator), by = c("ID", "condition")) %>%
  filter(administrator != "Martin/Oliver") %>%
  group_by(what, when) %>%
  group_modify(~broom::tidy(aov(.x$value ~ .x$administrator)), .keep = FALSE) %>%
  ungroup() %>%
  filter(term != "Residuals") %>%
  gt() %>%
    tab_header(title = "Effect of administrator") %>%
    fmt_number(columns = where(~ is.numeric(.x)), decimals = 3)

# ANOVA for vitality in post measurement by administrator
df_measures %>%
  filter(what == "vitality", when == "post") %>%
  left_join(select(df_forest, ID, condition, administrator), by = c("ID", "condition")) %>%
  filter(administrator != "Martin/Oliver") %>%
  aov(value ~ administrator, data = .) %>%
  summary()
```

```{r administrator effect plots}
df_forest %>%
  select(condition, administrator, contains("_diff")) %>%
  pivot_longer(-c(administrator, condition), names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = variable, y=value, fill=administrator, color=administrator)) +
    geom_boxplot() +
    geom_point(position=position_dodge(width=0.75)) +
    facet_grid(~condition) +
    coord_flip() +
    theme_minimal()
```

Testing if the fact that it is raining at the start of the experience has any effect on the outcome measures.

### Effect of weather

```{r rain effects}
head(df_forest)
create_post_table(df_measures, df_forest, "weather_rain")
check_difference(filter(df_measures, condition == "les"), df_forest, "weather_rain")
```

Interestingly there are no significant effects of rain on either of the measures.

```{r humidity effects}
plot_relationship(filter(df_measures, condition=="les"), df_forest, "humidity_avg")
check_correlations(filter(df_measures, condition=="les"), df_forest, "humidity_avg")
```

There seems to be no strong effect of humidity on the outcome measures.

```{r temperature effects}
plot_relationship(filter(df_measures, condition=="les"), df_forest, "temperature_avg")
check_correlations(filter(df_measures, condition == "les"), df_forest, "temperature_avg")
```

And neither does temperature seem to have any effect on the outcome measures.

### Heart rate in norm

```{r blood pressure table}
df_forest %>%
  select(condition, matches("blood_.*_pre")) %>%
  pivot_longer(-condition, names_to = "variable", values_to = "value") %>%
  group_by(condition, variable) %>%
  report_table()
```

```{r blood pressure norm plot}
df_forest %>%
  select(condition, matches("blood_.*_pre")) %>%
  pivot_longer(-condition, names_to = "variable", values_to = "value") %>%
  ggplot(aes(value)) +
    geom_histogram() +
    facet_grid(~variable)
```

### Correlation between blood measures and ROS and POMS
```{r blood pressure correlation}
df_forest %>%
  select(condition, ends_with("_diff")) %>%
  GGally::ggpairs(mapping = aes(color = condition))
```

## Effect of planes

```{r planes}
df_forest %>%
  select(ID, condition, airplanes, ends_with("post")) %>%
  pivot_longer(cols=-c(ID, condition, airplanes), names_to = "variable", values_to = "value") %>%
  ggplot(aes(airplanes, value)) +
    geom_point() +
    geom_smooth(method="lm") +
    facet_grid(rows=vars(variable), cols=vars(condition))
```

## Effect of sex

There doesnt seem to be any effect of gender or interaction between gender and efficiency.

```{r sex}
df_forest %>%
  select(ID, gender, condition, ends_with("diff")) %>%
  pivot_longer(cols=-c(ID, gender, condition), names_to = "variable", values_to = "value") %>%
  ggplot(aes(gender, value)) +
    geom_boxplot() +
    facet_grid(cols=vars(variable), rows=vars(condition), scales = "free_y")

check_difference(df_measures, df_forest, "gender")

filter(df_measures, condition == "les") %>%
  check_difference(df_forest, "gender")

filter(df_measures, condition == "VR") %>%
  check_difference(df_forest, "gender")
```