---
title: "Data checks"
author: "Lukáš Hejtmánek"
format: 
  html:
    self-contained: true
---

library(tidyverse)
library(tidyr)
library(here)

LOCAL_DATA_PATH <- "data"

source("functions/loading.R", encoding = "UTF-8")
source("functions/loading-questionnaire.R", encoding = "UTF-8")
source("functions/loading-apple.R", encoding = "UTF-8")
source("functions/processing-questionnaires.R", encoding = "UTF-8")

dir.create("temp", showWarnings = FALSE)

## Paper questionnaires -----
df_session1 <- load_paper_questionnaire_data("Session1")
df_session2 <- load_paper_questionnaire_data("Session2")

df_session1 <- mutate(df_session1, session = 1)
df_session2 <- mutate(df_session2, session = 2)

df_questionnaire <- bind_rows(df_session1, df_session2)
df_questionnaire <- add_relax_scores(df_questionnaire)
df_questionnaire <- add_poms_scores(df_questionnaire)

df_questionnaire <- add_blood_pressure_scores(df_questionnaire)
df_questionnaire$condition <- df_questionnaire$sessionType

df_questionnaire <- code_weather(df_questionnaire)

save(df_questionnaire, file = "temp/df_questionnaire.RData")

## ssq ----
df_ssq <- load_online_questionnaire_data("SSQ")
df_ssq <- process_ssq(df_ssq)
save(df_ssq, file = "temp/df_ssq.RData")

## quality ----
df_quality <- load_online_questionnaire_data("Quality")
df_quality <- process_quality(df_quality)
save(df_quality, file = "temp/df_quality.RData")

## immersion ----
df_immersion <- load_online_questionnaire_data("Immersion")
df_immersion <- process_immersion(df_immersion)
save(df_immersion, file = "temp/df_immersion.RData")

## BFI ---
# Needs explicitly state character
df_bfi <- read_sheet("1Cp9mc7GGKKMGybn63QLDlZctYChlda-3rHSiVudxgl8",
  sheet = "Qualtrics", col_types = "c", na = c("", "NA"))
df_bfi <- df_bfi[-1, ]
# remove second row with names of columns
df_bfi <- readr::type_convert(df_bfi, na = c("", "NA"))
df_bfi <- repsy.bfi::process_bfi(df_bfi, "Big5", "60")
df_bfi <- mutate(df_bfi, gender = recode(sex, `1` = "male", `2` = "female"))
#recode columnn gender in the tale df_bfi so that 2 is female and 1 is male
df_bfi <- standardize_bfi(df_bfi, "CZ60z")
df_bfi <- mutate(df_bfi, ID = toupper(ID))

## Labels for fields -----
field_labeller <- c("forestry.landscaping", "agriculture.horticulture",
  "botany.plant.biology", "biology.other", "environmental.science", "none")
## adds professions
df_bfi <- df_bfi %>%
  select("ID", starts_with("fieldProfession")) %>%
  pivot_longer(-ID) %>%
  separate(name, into = c("field", "category"), sep = "_") %>%
  mutate(category = field_labeller[as.numeric(category)]) %>%
  filter(!is.na(value)) %>%
  group_by(ID) %>%
  mutate(proffessions = paste0(category, collapse = ",")) %>%
  ungroup() %>%
  select(ID, proffessions) %>%
  distinct() %>%
  right_join(df_bfi, by = "ID") %>%
  select(-starts_with("fieldProfession"))

## Adds hobbies ----
df_bfi <- df_bfi %>%
  select("ID", starts_with("fieldHobby")) %>%
  pivot_longer(-ID) %>%
  separate(name, into = c("field", "category"), sep = "_") %>%
  mutate(category = field_labeller[as.numeric(category)]) %>%
  filter(!is.na(value)) %>%
  group_by(ID) %>%
  mutate(hobbies = paste0(category, collapse = ",")) %>%
  ungroup() %>%
  select(ID, hobbies) %>%
  distinct() %>%
  right_join(df_bfi, by = "ID") %>%
  select(-starts_with("fieldHobby"))

save(df_bfi, file = "temp/df_bfi.RData")

## Airplane data ----
df_airplanes <- load_airplane_data(file.path(LOCAL_DATA_PATH, "airplanes"))
df_airplanes

df_airplane_timestamps <-
  load_online_questionnaire_data("merging-timestamps") %>%
  filter(condition == "VR") %>%
  mutate(timestamp_hour = ifelse(
    is.na(airplane_hour_edit), timestamp_hour, airplane_hour_edit))

df_airplanes_count <- df_airplanes %>%
  group_by(date, hour) %>%
  summarise(airplanes = n()) %>%
  ungroup()

df_airplanes_sounds <- df_airplanes %>%
  left_join(select(df_airplane_timestamps, ID, condition,
    timestamp_hour, timestamp_date),
  by = c("hour" = "timestamp_hour", "date" = "timestamp_date")) %>%
  filter(!is.na(ID))

df_airplanes_count <- left_join(df_airplanes_count, df_airplane_timestamps,
  by = c("hour" = "timestamp_hour", "date" = "timestamp_date")) %>%
  select(ID, airplanes) %>%
  filter(!is.na(ID)) %>%
  mutate(condition = "VR")

save(df_airplanes_count, df_airplanes_sounds, file = "temp/df_airplanes.RData")

## Apple watch data ---
watch_data <- load_health_data(file.path(LOCAL_DATA_PATH, "apple_health_export/export.xml"))
df_health <- merge_health_workouts(watch_data$health, watch_data$workouts)

df_health <- df_health %>%
  filter(!is.na(workout_starttime)) %>%
  mutate(workout_date = str_glue("{year(workout_starttime)}-{month(workout_starttime)}-{day(workout_starttime)}"),
         workout_hour = str_glue("{hour(workout_starttime)}"))

df_watch_timestamps <-
  load_online_questionnaire_data("merging-timestamps", col_types = "c") %>%
  mutate(timestamp_hour = ifelse(
    is.na(watch_hour_edit), timestamp_hour, watch_hour_edit))

df_health <- df_health %>%
  left_join(select(df_watch_timestamps, ID, condition, timestamp_hour, timestamp_date),
    by = c("workout_hour" = "timestamp_hour", "workout_date" = "timestamp_date")) %>%
  filter(!is.na(ID))

save(df_health, file = "temp/df_health.RData")
rm(watch_data)
## Merging ----
df_forest <- df_questionnaire %>%
  left_join(df_ssq, by = "ID") %>%
  left_join(df_quality, by = "ID") %>%
  left_join(df_immersion, by = "ID") %>%
  left_join(df_bfi, by = "ID") %>%
  rows_update(df_airplanes_count, by = c("ID", "condition"))

save(df_forest, file = "temp/df_forest.RData")



## Data checks

```{r}

```

## Dropping participants

FOVR 24 last questionnaire (completely last person) was lost.

### Apple processing

```{r apple watch processing}

```

FOVR 33 forest is missing (10.11.2023 15:00 - forgot to start?)
FOVR 24
First 5 participants are missing as well (forest Data - forgot the watch).

